SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-12-23 maintex.tex
more text revisions

Once the geometries of the galaxies are set, we can build azimuthally averaged radial profiles. The spaxels are binned into radius increments of 0.2 \reff\ from 0.0$-$2.6 \reff. \textbf{The MaNGA sample does not have a uniform spatial coverage, 63\% of the sample is designed to cover 1.5 \reff\ (the Primary+ subsample) and 37\% of the sample is designed to cover 2.5 \reff\ (the Secondary subsample).} While the MaNGA sample is limited spatially to 1.5$-$2.5 \reff\, the spaxels along the semi-minor axis of a galaxy will be able to extend beyond that range because of the galaxy's inclination angle, sin($i$).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{fig/ssfr_mass.pdf}
\caption[]{The \textbf{Left} panel shows the $\Delta$log(sSFR) profiles where the difference profiles are constructed from the difference between the paired galaxy profiles and a set of 20 control galaxies. The profiles are split into five different stellar mass bins and the highlighted region about the profiles represent the standard error of mean of the profile. The black dashed line represents the mean profile between the four difference mass ranges. The number of paired galaxies in each mass range is given in the legend in parentheses. In the \textbf{Right} panel the nuclear $\Delta$log(sSFR) are shown. The black squares are the mean values within a stellar mass bin (where the size of the bins are shown the the horizontal error bars). The vertical error bars on the black squares represent the standard deviation within the bin. The horizontal, dashed black line represents the median central enhancement of the pair sample. Galaxies in the outside-IFU (Blue) and inside-IFU (Red) samples are separately depicted.}
\label{fig:ssfr_mass}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The mass ratio here is defined as, 
\begin{equation}
\Delta {\rm log}(M) = {\rm log}(M_{\rm target}) - {\rm log}(M_{\rm comp}),
\end{equation}
where $M_{\rm target}$ is the stellar mass of the MaNGA target galaxy and $M_{\rm comp}$ is the stellar mass of the companion galaxy. In the sample, the MaNGA target is not always the more massive component of the pair. We split the sample into two groups based on their mass ratios; primary companions in which the MaNGA target is the more massive galaxy and secondary companions in which the MaNGA target is the less massive galaxy.

@@

In the top left of Figure \ref{fig:ssfr_dm}, we split the profiles by mass ratio in four bins; the primary and secondary companion of a major merger, $|\Delta$log($M$)$|$ $<$ 0.5, and the primary and secondary companion of a minor merger, $|\Delta$log($M$)$|$ $>$ 0.5. Regardless of the mass ratio, the centers of the paired galaxies are centrally enhanced, though major mergers have stronger enhancements in their centers by $\sim$0.2 dex over minor mergers. Secondary companions also feature slightly higher levels of enhancement in their centers than primary companions at the same mass ratios by $\sim$0.1 dex.

@@

In the top right of Figure \ref{fig:ssfr_dm} we see that the sSFR enhancement is strongest in galaxies close to a 1:1 mass ratio ($\Delta$log($M$) $=$ 0.0). These galaxies feature a central enhancement of $\sim$0.35 dex which is 0.1 dex higher than the average central enhancement of the whole sample. This enhancement falls with wider mass ratio; however, pairs with large differences in stellar mass still feature substantial levels of star formation enhancement, 0.15$-$0.20 dex at mass ratios of $|\Delta$log($M$)$|$ $=$ 1.0. We also see that secondary companions show a higher level of sSFR enhancement, by $\sim$0.1 dex, with respect to primary companions.

@@

Figure \ref{fig:ssfr_sep} shows the profile and the central level of sSFR enhancement as a function of the projected separation ($r_{\rm p}$) from our MaNGA data. The $\Delta$log(sSFR) profiles show a clear gradient with the projected separation with close galaxy pairs showing higher levels of sSFR enhancement compared to pairs with wide separations. The profiles in the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 0.0$-$12.5, are generally $\sim$0.1$-$0.15 dex above the median and the profiles of the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 37.5$-$50.0, are generally $\sim$0.05$-$0.1 dex below the median.

>>>>

Once the geometries of the galaxies are set, we can build azimuthally averaged radial profiles. The spaxels are binned into radius increments of 0.2 \reff\ from 0.0$-$2.6 \reff. \textbf{The MaNGA sample does not have a uniform spatial coverage, 63\% of the sample is designed to cover 1.5 \reff\ (the Primary+ subsample) and 37\% of the sample is designed to cover 2.5 \reff\ (the Secondary subsample).} {\it While the MaNGA sample is limited spatially to 1.5$-$2.5 \reff\, the spaxels along the semi-minor axis of a galaxy will be able to extend beyond that range because of the galaxy's inclination angle, sin($i$). }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\centering
\includegraphics[width=0.8\linewidth]{fig/ssfr_mass.pdf}
\caption[]{The \textbf{Left} panel shows the $\Delta$log(sSFR) profiles where the difference profiles are constructed from the difference between the paired galaxy profiles and a set of 20 control galaxies. The profiles are split into five different stellar mass bins and the highlighted region about the profiles represent the standard error of mean of the profile. The black dashed line represents the mean profile between the four difference mass ranges. The number of paired galaxies in each mass range is given in the legend in parentheses. The \textbf{Right} panel shows the nuclear $\Delta$log(sSFR) \textbf{extracted from a 0.5\reff\ aperture}. The black squares are the mean values within a stellar mass bin (where the size of the bins are shown the the horizontal error bars). The vertical error bars on the black squares represent the standard deviation within the bin. The horizontal, dashed black line represents the median central enhancement of the pair sample. Galaxies in the outside-IFU (Blue) and inside-IFU (Red) samples are separately depicted.}
\label{fig:ssfr_mass}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The mass ratio here is defined as, 
\begin{equation}
\Delta {\rm log}(M) = {\rm log}(M_{\rm target}) - {\rm log}(M_{\rm comp}),
\end{equation}
where $M_{\rm target}$ is the stellar mass of the MaNGA target galaxy and $M_{\rm comp}$ is the stellar mass of \textbf{its identified} companion galaxy. \textbf{We not only use the mass ratio to study how a major merger behaves differently from a minor merger but also to study how the more massive galaxy of a pair behaves differently from the less massive galaxy of the pair. We expect that either component of a similar mass pair will behave in a similar fashion; however, in a minor merger the large central galaxy may not respond to the merger event in the same way than its companion dwarf galaxy. The two different scenarios can be differentiated with our definition of the mass ratio; when the MaNGA target is the more massive companion of the pair, its mass ratio is positive and when MaNGA target is the less massive companion of pair the mass ratio is negative.}

%{\it In the sample, the MaNGA target is not always the more massive component of the pair. We split the sample into two groups based on their mass ratios; primary companions in which the MaNGA target is the more massive galaxy and secondary companions in which the MaNGA target is the less massive galaxy.}

@@

In \textbf{left panel} of Figure \ref{fig:ssfr_dm}, we split the profiles by mass ratio in four bins; the primary and secondary companion of a major merger, $|\Delta$log($M$)$|$ $<$ 0.5, and the primary and secondary companion of a minor merger, $|\Delta$log($M$)$|$ $>$ 0.5. Regardless of the mass ratio, the centers of the paired galaxies are centrally enhanced, though major mergers have stronger enhancements in their centers by $\sim$0.2 dex over minor mergers. Secondary companions also feature slightly higher levels of enhancement in their centers than primary companions at the same mass ratios by $\sim$0.1 dex.

@@

In \textbf{right panel} of Figure \ref{fig:ssfr_dm} we see that the sSFR enhancement is strongest in galaxies close to a 1:1 mass ratio ($\Delta$log($M$) $=$ 0.0). These galaxies feature a central enhancement of $\sim$0.35 dex which is 0.1 dex higher than the average central enhancement of the whole sample. This enhancement falls with wider mass ratio; however, pairs with large differences in stellar mass still feature substantial levels of star formation enhancement, 0.15$-$0.20 dex at mass ratios of $|\Delta$log($M$)$|$ $=$ 1.0. We also see that secondary companions show a higher level of sSFR enhancement, by $\sim$0.1 dex, with respect to primary companions.

@@

Figure \ref{fig:ssfr_sep} shows the profile and the central level of sSFR enhancement as a function of the projected separation ($r_{\rm p}$) from our MaNGA data. \textbf{The $\Delta$log(sSFR) profiles show only a weak dependency on the projected separation. The pairs below a separation of 12.5 kpc have a sSFR enhancement which is roughly 0.1 dex above pairs with wider separations within 0.2 \reff. }

%{\it The $\Delta$log(sSFR) profiles show a clear gradient with the projected separation with close galaxy pairs showing higher levels of sSFR enhancement compared to pairs with wide separations. The profiles in the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 0.0$-$12.5, are generally $\sim$0.1$-$0.15 dex above the median and the profiles of the galaxy pairs in the separation range, $r_{\rm p}$ $=$ 37.5$-$50.0, are generally $\sim$0.05$-$0.1 dex below the median. }
